Extended Data

### 2021-11-16 doc/manuscript/Dow_manuscript_Nature.Rmd
tweak abstract

**As the climate changes, warmer spring temperatures are causing earlier leaf-out[@jeong_phenology_2011;@keenan_net_2014; @buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020] and commencement of net carbon dioxide (CO~2~) sequestration[@keenan_net_2014;@finzi_carbon_2020] in temperate deciduous forests, resulting in a tendency towards increased growing season length[@menzel_growing_1999;@menzel_european_2006;@ibanez_forecasting_2010;@jeong_phenology_2011;@finzi_carbon_2020;@stuble_plant_2021] and annual CO~2~ uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014; @crabbe_extreme_2016; @zhou_explaining_2016; @fu_maximum_2019;@finzi_carbon_2020].**
**However, less is known about how spring temperatures affect tree stem growth, which sequesters carbon in wood that has a long residence time in the ecosystem [@xue_global_2017;@russell_residence_2014].**
**Using dendrometer band measurements from `r n_dendro_trees` trees across two forests, we show that warmer spring temperatures consistently shifted the woody growth of deciduous trees earlier but had no consistent effect on peak growing season length, maximum daily growth rates, or annual growth.**
**The latter finding was confirmed on the centennial scale by `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America, where annual growth was far more sensitive to temperatures during the peak growing season than in the spring.**
**These findings imply that extra CO~2~ sequestered in years with warmer springs[@keeling_increased_1996;@dragoni_evidence_2011;@crabbe_extreme_2016] is not allocated to long-lived woody biomass, where it could have a substantial and lasting impact on the forest C balance. **
**Such empirical results contradict current projections from global carbon cycle models[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012], and imply that warming spring temperatures are unlikely to increase the woody productivity or strengthen the carbon sink of temperate deciduous forests.**

@@

Here, we evaluate how early spring temperatures affect stem growth phenology, growth rates, and annual growth of temperate deciduous trees in two forests and annual growth across forests spanning eastern North America. 
To test whether warmer springs extend the period of stem growth, we used dendrometer band measurements on r n_dendro_trees` trees across two deciduous forests. 
To test whether spring temperatures consistently increased annual growth, we used `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America.

>>>>

**As the climate changes, warmer spring temperatures are causing earlier leaf-out[@jeong_phenology_2011;@keenan_net_2014; @buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020] and commencement of net carbon dioxide (CO~2~) sequestration[@keenan_net_2014;@finzi_carbon_2020] in temperate deciduous forests, resulting in a tendency towards increased growing season length[@menzel_growing_1999;@menzel_european_2006;@ibanez_forecasting_2010;@jeong_phenology_2011;@finzi_carbon_2020;@stuble_plant_2021] and annual CO~2~ uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014; @crabbe_extreme_2016; @zhou_explaining_2016; @fu_maximum_2019;@finzi_carbon_2020].**
**However, less is known about how spring temperatures affect tree stem growth, which sequesters carbon in wood that has a long residence time in the ecosystem[@xue_global_2017;@russell_residence_2014].**
**Using dendrometer band measurements from `r n_dendro_trees` trees across two forests, we show that warmer spring temperatures consistently shifted the woody growth of deciduous trees earlier but had no consistent effect on peak growing season length, maximum daily growth rates, or annual growth.**
**The latter finding was confirmed on the centennial scale by `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America, where annual growth was far more sensitive to temperatures during the peak growing season than in the spring.**
**These findings imply that extra CO~2~ uptake in years with warmer springs[@keeling_increased_1996;@dragoni_evidence_2011;@crabbe_extreme_2016] is not allocated to long-lived woody biomass, where it could have a substantial and lasting impact on the forest C balance. **
**Rather, contradicting current projections from global carbon cycle models[@buermann_widespread_2018; @zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012], our empirical results imply that warming spring temperatures are unlikely to increase the woody productivity or strengthen the carbon sink of temperate deciduous forests.**

@@

Here, we evaluate how early spring temperatures affect stem growth phenology, growth rates, and annual growth of temperate deciduous trees in two forests and annual growth across forests spanning eastern North America. 
To test whether warmer springs extend the period of stem growth, we used dendrometer band measurements on r n_dendro_trees` trees across two mid-latitude deciduous forests. 
To test whether spring temperatures consistently increased annual growth, we used `r n_chronologies` tree-ring chronologies from `r n_cores_sites` forests across eastern North America.
